---
title: "Assignment-4"
author: "TarakRam Nunna"
date: "05/11/2021"
output: html_document
---

## setting working directory
```{r}
getwd()
setwd("C:/Users/TARAKRAM/OneDrive/Desktop/ML_code/64060_tnunna/Assignment-4")
Pharmaceuticaldata<- read.csv("Pharmaceuticals.csv")
str(Pharmaceuticaldata)
```

## installing required packages
```{r}
library(tidyverse)
library(factoextra)
library(cluster)
library(ggplot2)
library(gridExtra)
```

# a. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 

# remove any missing value(NA) that might be present in the Pharmaceuticaldata and selecting numerical data
```{r}
Pharmaceuticaldata <- na.omit(Pharmaceuticaldata)
Pharmaceuticaldata
row.names(Pharmaceuticaldata)<- Pharmaceuticaldata[,1]
Pharmadata<- Pharmaceuticaldata[, 3:11]
head(Pharmadata)
```

# Scaling the Pharmadata using the scale function
```{r}
Pharmascale <- scale(Pharmadata)
head(Pharmascale)
```
# Computing K-means clustering in R for different centers 
# Using multiple values of K and examine the differences in results
```{r}
kmeans1 <- kmeans(Pharmascale, centers = 2, nstart = 30)
kmeans2<- kmeans(Pharmascale, centers = 5, nstart = 30)
kmeans3<- kmeans(Pharmascale, centers = 6, nstart = 30)
Plot1<-fviz_cluster(kmeans1, data = Pharmascale)+ggtitle("k=2")
plot2<-fviz_cluster(kmeans2, data = Pharmascale)+ggtitle("k=5")
plot3<-fviz_cluster(kmeans3, data = Pharmascale)+ggtitle("k=6")
grid.arrange(Plot1,plot2,plot3, nrow = 3)
```

```{r}
distance<- dist(Pharmascale, method = "euclidean")
fviz_dist(distance)
```

# For each k, calculate the total within-cluster sum of square (wss)
# Compute and plot wss for k = 1 to k = 10
# extract wss for 2-15 clusters
# The location of a bend (knee) in the plot is generally considered as an indicator of the appropriate number of clusters k =5.
```{r}
set.seed(300)
wss<- function(k){kmeans(Pharmascale, k, nstart =10)$tot.withinss}
k.values<- 1:10 
wss_clusters<- map_dbl(k.values, wss)
plot(k.values, wss_clusters, type="b", pch = 16, frame = TRUE, xlab="Clusters", ylab="WSS")
```

# Final analysis and Extracting results using 5 clusters and Visualize the results
```{r}
set.seed(300)
final_Cluster<- kmeans(Pharmascale, 5, nstart = 25)
print(final_Cluster)
clusplot(Pharmascale,final_Cluster$cluster, color = TRUE, labels = 2,lines = 0)
```

b) Interpret the clusters with respect to the numerical variables used in forming the clusters
Cluster 1 - AHM,SGP,WYE,BMY,AZN, ABT, NVS, LLY
Cluster 2 - BAY, CHTT, IVX
Cluster 3 - WPI, MRX,ELN,AVE
Cluster 4 - AGN, PHA 
Cluster 5 - JNJ, MRK, PFE,GSK

```{r}
Pharma_Cluster <- Pharmaceuticaldata[,c(12,13,14)]%>% mutate(clusters = final_Cluster$cluster)%>% arrange(clusters, ascending = TRUE)
Pharma_Cluster
```

# c)Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
```{r}
plot1<-ggplot(Pharma_Cluster, mapping = aes(factor(clusters), fill=Median_Recommendation))+geom_bar(position = 'dodge')+labs(x ='No of clusters')
plot2<- ggplot(Pharma_Cluster, mapping = aes(factor(clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='No of clusters')
plot3<- ggplot(Pharma_Cluster, mapping = aes(factor(clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='No of clusters')
grid.arrange(plot1, plot2, plot3)       
```

As per graph, 
Cluster 1 Suggests to Hold to Moderate Buy to Moderate Sell
Cluster 2 Suggests to Hold to Moderate Buy
Cluster 3 Suggests to Moderate Buy to Moderate Buy
Cluster 4 suggests to Hold to Moderate Buy
Cluster 5 suggests to Hold to Moderate Buy

# d)Provide an appropriate name for each cluster using any or all of the variables in the dataset.
Cluster1-Hold Cluster
Cluster2-Hold Cluster
Cluster3-Buy-Sell Cluster
Cluster4-Hold-Buy Cluster
Cluster5-Hold-Buy Cluster
